---
title: "Occupancy dynamic models - Sandhill crane"
author: "Clara Casbona"
date: "28 Jul 2021"
output:
  pdf_document: default
  word_document: default
---

This code cover the basic steps for applying occupancy dynamic models to the  sandhill crane.

Period : April - June, 2004 - 2019

Study area : South 51 parallel, Quebec

Data : Simulated data 

# Packages : 

```{r Directory, include=FALSE}
library(AICcmodavg)
library(unmarked)
library(dplyr)
library(ggplot2)
library(ggcorrplot)
library(tidymv)
library(ggpubr)
library(reshape2)
library(extrafont)
```

# Loading data:

```{r Directory2, include=FALSE}
grue_dyn <- read.csv("../data/Simulated_data.csv", header = TRUE) #Database

# If you don't want to run everything, You can load this data: 

load("../data/MHo_p_MAt.sMAt2_MHo_p.Rdata") # To make the run faster
#load("../data/Mod_long.Rdata")

```

Split data-set (detection (1/0), observation covariables and site covariables)

```{r Directory3, include=FALSE}

# Prepare tables for unmarked: 
det   <- grue_dyn[,c(2:97)]
Met   <- grue_dyn[,c(102:197)] 
Met_eff <-  grue_dyn[,c(198:293)] 
Eff   <- grue_dyn[,c(390:485)]
JJ    <- grue_dyn[,c(486:581)]
SiteCovs <- grue_dyn[,c(582:594)]
YearCovs_MAt <- grue_dyn[,c(595:610)]
YearCovs_MAi <- grue_dyn[,c(611:626)]

## set up yearly site covariate
year <- matrix(data = c( 2004, 2005, 2006, 2007, 2008, 2009, 2010, 2011, 2012, 2013, 2014, 2015, 2016, 2017, 2018, 2019),
               nrow = nrow(JJ), 16, byrow = TRUE)


colnames(SiteCovs) <- c("FOBO_p", "EPe_p", "EPf_p", "EPl_p", "EPm_p", "MHo_p", "mm", "temp", "MU_p", " km", "LONG", "MAt_s", "MAi_s")
```

Standardising covariates: 

```{r Standardisation - Preparation pour unmarked, message=FALSE}

# Site variables 

SiteCovs_s <- SiteCovs %>% mutate_each_(funs(scale(.) %>% as.vector), vars=c( "MHo_p","MU_p", "mm", "temp", "LONG", "MAi_s", "MAt_s"))

SiteCovs_s$MAt_2 <- SiteCovs_s$MAt_s^2 
SiteCovs_s$MAi_2 <- SiteCovs_s$MAi_s^2 

# Yearly changing variables (habitat + observational variables )
MAi <- as.matrix(YearCovs_MAi)
MAi.mean <- mean(MAi, na.rm = TRUE) 
MAi.sd <- sd(MAi, na.rm = TRUE)
MAi.s <- (MAi-MAi.mean)/MAi.sd
MAi2 <- MAi.s^2 

MAt <- as.matrix(YearCovs_MAt)
MAt.mean <- mean(MAt, na.rm = TRUE) 
MAt.sd <- sd(MAt, na.rm = TRUE)
MAt.s <- (MAt-MAt.mean)/MAt.sd
MAt2 <- MAt.s^2 

Eff <- as.matrix(Eff)
Eff.mean <- mean(Eff, na.rm = TRUE) 
Eff.sd <- sd(Eff, na.rm = TRUE)
Eff.std <- (Eff-Eff.mean)/Eff.sd
Eff2 <- Eff.std^2


JJ <- as.matrix(JJ)
JJ.mean <- mean(JJ, na.rm = TRUE) 
JJ.sd <- sd(JJ, na.rm = TRUE)
JJ.std <- (JJ-JJ.mean)/JJ.sd
JJ2 <- JJ.std^2

year <- as.matrix(year)
year.mean <- mean(year, na.rm = TRUE) 
year.sd <- sd(year, na.rm = TRUE)
year.std <- (year-year.mean)/year.sd



```

We convert this data into an unmarked object so we can start fitting occupancy models.

```{r Formatage des donnees pour unmarked, echo=FALSE}
## Function
grue.data <- unmarkedMultFrame(y = det, 
                                siteCovs = SiteCovs_s,
                                obsCovs = 
                                 list(Met = Met, 
                                  Met_eff = Met_eff,
                                  Eff = Eff.std,
                                  JJ = JJ.std),
                                yearlySiteCovs = 
                                list(year = year.std,
                                     MAi.s = MAi.s,
                                     MAt.s = MAt.s,
                                     MAi2 = MAi2,
                                     MAt2 = MAt2), 
                                numPrimary = 16)
## summary for grue.data 
summary(grue.data)

detHist(grue.data)


# DETECTION
hist <- detHist(grue.data)
hist <- hist$out.freqs
str(hist)

hist.df=data.frame(hist)
season <- factor(c(2004:2019))
hist.df <- cbind(hist.df, season)

write.csv(hist.df, file = "../output/histdf.csv")

hist.df1 <- hist.df[,c(1,7)]
hist.df2 <- hist.df[,c(2:4,7)]

his.melt=melt(hist.df1,var.id=season)


samplet <- ggplot(data=his.melt, aes(x=season, y=value,group=variable)) +
  geom_line(aes(color=variable), size = 0.5,color='darkblue')+
  geom_point(aes(color=variable, shape=variable), size = 2,color='darkblue')+
  xlab("") + ylab("") +
    ggtitle("1A") +
  theme_classic()+
  theme(plot.title = element_text(size=15), 
        axis.text=element_text(size=15),
        axis.title=element_text(size=15,face="bold"),
        axis.text.x = element_text(angle = 45,hjust = 1),
        legend.position = c(0.1, 0.8),
        legend.title = element_text(size =15)) 

his.melt=melt(hist.df2,var.id=season)


histogr <-ggplot(data=his.melt, aes(x=season, y=value, group=variable)) +
  geom_line(aes(color=variable), size = 0.5)+
  geom_point(aes(color=variable,shape=variable), size = 2)+
  xlab("") + ylab("") +
theme_classic()+
    ggtitle("1B") +
  scale_colour_brewer(palette = "Set1")+
  theme(plot.title = element_text(size=15), 
        axis.text=element_text(size=15),
        axis.title=element_text(size=15,face="bold"),
        axis.text.x = element_text(angle = 45,hjust = 1),
        legend.position = c(0.1, 0.8),
        legend.title = element_text(size =15))

#Unification des deux graphiques dans un

figure <- ggarrange(samplet, histogr + font("x.text", size = 15),
                    ncol = 1, nrow = 2)
annotate_figure(figure,
                top = text_grob(""),
                bottom = text_grob("Year", size = 15),
                left = text_grob("Number of cells", rot = 90))


```

Verifying correlations

```{r exploration inital et correlations}
# Visualisation des donnes SiteCovs


corr<- cor(SiteCovs[,6:11]) # aucune correlation > 0.4
corr <- round(corr, 2)
# Visualize correlogram  for site obs
ggcorrplot(corr, sig.level=0.05, lab_size = 4.5, p.mat = NULL, 
           lab = TRUE,color = c("#FC4E07", "white", "#00AFBB"),
           insig = c("pch", "blank"), pch = 1, pch.col = "black", pch.cex =1,
           tl.cex = 12) +
  theme(axis.text.x = element_text(margin=margin(-2,0,0,0)),  # Order: top, right, bottom, left
        axis.text.y = element_text(margin=margin(0,-2,0,0))) +
  geom_vline(xintercept=1:ncol(corr)-0.5, colour="white", size=2) +
  geom_hline(yintercept=1:ncol(corr)-0.5, colour="white", size=2) 


```


Now that we've created the data frame with detection histories and covariates, we use unmarked to fit a multiple-season occupancy model - MacKenzie et al. (2017). 

We  fit the occupancy model to our data using the coltex() function, specifying the occupancy, colonisation, extinction and detection covariates, respectively. 

Step 1: 

First we created a list of models using covariantes for detection and selecting the best one to build the complet models.

```{r Liste des modeles detection, echo=TRUE}
# Liste de mod?les pour la d?tection 
m0 <-  colext(~ 1, ~ 1, ~ 1, ~ JJ, data = grue.data)
m1 <-  colext(~ 1, ~ 1, ~ 1, ~ JJ + Met_eff-1, data = grue.data)
m2 <-  colext( ~ 1,  ~ 1, ~ 1, ~ JJ + Eff,  data = grue.data)
m3 <- colext(psiformula = ~ 1, ~ 1, ~ 1, ~ JJ + Met -1, data = grue.data)

```

# AICc selection

```{r Selection AIC, echo=FALSE}

# Selection AICcmodavg

Candidates <- list(m0, m1, m2, m3)
Names <- c("m0", "m1", "m2", "m3")
aictab(cand.set = Candidates, modnames = Names, c.hat = 1)
save(m1, file= "Detect_model.Rdata")
# check parametres
evidence(aictab(cand.set = Candidates, c.hat = 1))
lapply(Candidates, extractCN)
lapply(Candidates, checkParms, simplify = FALSE)

# checker convergence 
lapply(Candidates, checkConv) 

```


Test GOF (Runing this take several hours) - You have to activated to run using eval = TRUE, include = TRUE
```{r, eval=FALSE, include=FALSE}
#Test de MacKenzie et Bailey for checking model adjustment

gof <- mb.gof.test(m1, nsim = 1000, plot.hist = TRUE) 
save(gof, file = "m1_dyn_gof.Rdata")


## C.hat

gof$chisq.table
gof$c.hat.est


# Plot
hist(gof$t.star, 
     main = as.expression(substitute("Bootstrapped "*chi^2*" fit statistic ("*nsim*" samples)",
         list(nsim = gof$nsim))), 
     xlim = range(c(gof$t.star, gof$chi.square)), 
     xlab = paste("Simulated statistic ", 
       "(observed = ", round(gof$chi.square, digits = 2), ")", 
       sep = ""))
title(main = bquote(paste(italic(P), " = ", .(gof$p.global))), 
      line = -0.5)
abline(v = gof$chi.square, lty = "dashed", col = "red")



```


Detection figure

```{r}
## Detection data
nd <- data.frame(Met_eff=c('ATL_E','ATL_F','EBIRD_E','EBIRD_F','HHT'), JJ = 0)
E.det <- predict( m1, type='det', newdata=nd)

save(E.det, file = "../output/E.det.Rdata")

with(E.det, {
  plot(1:5, Predicted, pch=19, xaxt='n', xlab='Sampling methodology split by effort',
       ylab=expression(paste('Detection probability ( ', p, ' )')),
       ylim=c(0,1), col= c("black", "black", "black", "black", "black"),
       xlim = c(0.5,5.5), 
       cex.axis = .9,
       cex.lab = .8)
  axis(1, at=1:5, labels=c("Atlas high", "Atlas low","eBird high", "eBird low", "CWS"),cex.axis = .7 )
  arrows(1:5, lower, 1:5, upper, code=3, angle=90, length=0.05, col=  c("black", "black", "black", "black", "black"))
})

```
Step 2.1:  Expansion models

```{r modeles dynamiques LONG, echo=FALSE}
MOD.LONG <- colext( ~ LONG , ~ LONG , ~ LONG , ~ JJ + Met_eff , grue.data )
LONGYEAR <- colext( ~ 1 , ~ LONG + year + LONG*year , ~ 1  , ~ JJ + Met_eff , grue.data )
LONGYEAR2 <- colext( ~ LONG , ~ LONG + year + LONG*year , ~ LONG + year + LONG*year  , ~ JJ + Met_eff , grue.data )

save(MOD.LONG,LONGYEAR,LONGYEAR2, file = "Mod_long.Rdata")
load("Mod_long.Rdata")
summary(MOD.LONG)
summary(LONGYEAR2)

modYEAR2 <- colext( ~1 , ~ year  , ~ year   , ~ JJ + Met_eff , grue.data )
```

Expansion models figures

```{r figures LONG, echo=FALSE, warning=FALSE}

# Expansion prediction plots

LONG.mean <- mean(SiteCovs$LONG)
LONG.sd <- sd(SiteCovs$LONG)

year_vector <- c (-1.6269476, -1.4100213, -1.1930949, -0.9761686, -0.7592422, -0.5423159, -0.3253895, -0.1084632,  0.1084632,  0.3253895, 0.5423159,  0.7592422,  0.9761686,  1.1930949,  1.4100213,  1.6269476)

colonization_figures <- list()
extinction_figures   <- list()
  
for (i in 1:length(year_vector)) {
  
 New_data <- data.frame(LONG.t = seq(from=min(-80, na.rm = TRUE),
                                     to = max(-69, na.rm = TRUE),
                                     length=3600),
                        JJ = 0,     
                        year = year_vector[i] ,
                        Met_eff = factor("HHT",levels =levels(grue.data@obsCovs$Met_eff)))

 New_data$LONG <- (New_data$LONG.t-LONG.mean)/LONG.sd 

 
 # Colonization 
 
 E.col <- predict(LONGYEAR2, type="col", newdata=New_data, appendData=TRUE)
 E.col$Year <- i + 2003
 E.col$fit <- E.col$Predicted
 E.col$se.fit <- E.col$SE
 E.col$lwl <- E.col$fit-1.96*E.col$se.fit
 E.col$upl <- E.col$fit+1.96*E.col$se.fit
 
 Plot_colonization <- E.col %>%
  ggplot(aes(LONG.t, fit)) +
  geom_smooth_ci(color = "black") +
  theme_bw() +
  ylim(0,1.02) +
  geom_line(linetype = "dotted", aes(y = lwl), color = "grey43") +
  geom_line(linetype = "dotted", aes(y = upl), color = "grey43") +
  xlab("Longitude") + 
   annotate(family="Times New Roman", geom = "text", y =0.9 ,  x= -78, label = i + 2003, size = 8) +
  ylab(expression(paste('Colonization probability ( ', gamma, ' )'))) +
  theme(
      axis.title.x = element_text(family="Times New Roman", color="black", size = 16),
      axis.title.y = element_text(family="Times New Roman",color="black", size=16),
      axis.text=element_text(family="Times New Roman",size=12)
    )
 
 
  colonization_figures[[i]] <- Plot_colonization # Save all figures
  
 # Extinction 
 E.ext <- predict(LONGYEAR2, type="ext", newdata=New_data, appendData=TRUE)
 E.ext$Year <- i + 2003
 E.ext$fit <- E.ext$Predicted
 E.ext$se.fit <- E.ext$SE
 E.ext$lwl <- E.ext$fit-1.96*E.ext$se.fit
 E.ext$upl <- E.ext$fit+1.96*E.ext$se.fit
 
 Plot_extinction <- E.ext %>%
  ggplot(aes(LONG.t, fit)) +
  geom_smooth_ci(color = "black") +
  theme_bw() +
  ylim(-0.02,1) +
  geom_line(linetype = "dotted", aes(y = lwl), color = "grey43") +
  geom_line(linetype = "dotted", aes(y = upl), color = "grey43") +
  xlab("Longitude") + 
  annotate(family="Times New Roman", geom = "text", y =0.9 ,  x= -78, label = i + 2003, size = 8) +
  ylab(expression(paste('Extinction probability ( ', epsilon, ' )'))) +
  theme(
      axis.title.x = element_text(family="Times New Roman", color="black", size = 16),
      axis.title.y = element_text(family="Times New Roman",color="black", size=16),
      axis.text=element_text(family="Times New Roman",size=12)
    )
  extinction_figures[[i]] <- Plot_extinction # Save all figures

}

gridExtra::grid.arrange(grobs = colonization_figures[c(1,9,16)])
gridExtra::grid.arrange(grobs = extinction_figures[c(1,9,16)])


```

Step 2.3: Habitat models 

Running habitat models (This might take 1-2 hours)- You have to activated to run using eval = TRUE, include = TRUE

```{r modeles dynamiques habitat - H1, eval=FALSE, include=FALSE}
# List of 64 habitat models used in the AICc selection

MHo_p_MHo_p_MHo_p<-colext(~MHo_p,~MHo_p,~MHo_p,~JJ+Met_eff,grue.data)
MU_p_MHo_p_MHo_p<-colext(~MU_p,~MHo_p,~MHo_p,~JJ+Met_eff,grue.data)
MAt_sMAt_2_MHo_p_MHo_p<-colext(~MAt_s+MAt_2,~MHo_p,~MHo_p,~JJ+Met_eff,grue.data)
MAi_sMAi_2_MHo_p_MHo_p<-colext(~MAi_s+MAi_2,~MHo_p,~MHo_p,~JJ+Met_eff,grue.data)
MHo_p_MU_p_MHo_p<-colext(~MHo_p,~MU_p,~MHo_p,~JJ+Met_eff,grue.data)
MU_p_MU_p_MHo_p<-colext(~MU_p,~MU_p,~MHo_p,~JJ+Met_eff,grue.data)
MAt_sMAt_2_MU_p_MHo_p<-colext(~MAt_s+MAt_2,~MU_p,~MHo_p,~JJ+Met_eff,grue.data)
MAi_sMAi_2_MU_p_MHo_p<-colext(~MAi_s+MAi_2,~MU_p,~MHo_p,~JJ+Met_eff,grue.data)

# This was the top ranked model in the AICc selection using our original data - set

MHo_p_MAt.sMAt2_MHo_p <-colext(~MHo_p,~MAt.s+MAt2,~MHo_p,~JJ+Met_eff,grue.data) 

#save(MHo_p_MAt.sMAt2_MHo_p, file = "../output/MHo_p_MAt.sMAt2_MHo_p.Rdata")


MU_p_MAt.sMAt2_MHo_p<-colext(~MU_p,~MAt.s+MAt2,~MHo_p,~JJ+Met_eff,grue.data)
MAt_sMAt_2_MAt.sMAt2_MHo_p<-colext(~MAt_s+MAt_2,~MAt.s+MAt2,~MHo_p,~JJ+Met_eff,grue.data)
MAi_sMAi_2_MAt.sMAt2_MHo_p<-colext(~MAi_s+MAi_2,~MAt.s+MAt2,~MHo_p,~JJ+Met_eff,grue.data)
MHo_p_MAi.sMAi2_MHo_p<-colext(~MHo_p,~MAi.s+MAi2,~MHo_p,~JJ+Met_eff,grue.data)
MU_p_MAi.sMAi2_MHo_p<-colext(~MU_p,~MAi.s+MAi2,~MHo_p,~JJ+Met_eff,grue.data)
MAt_sMAt_2_MAi.sMAi2_MHo_p<-colext(~MAt_s+MAt_2,~MAi.s+MAi2,~MHo_p,~JJ+Met_eff,grue.data)
MAi_sMAi_2_MAi.sMAi2_MHo_p<-colext(~MAi_s+MAi_2,~MAi.s+MAi2,~MHo_p,~JJ+Met_eff,grue.data)
MHo_p_MHo_p_MU_p<-colext(~MHo_p,~MHo_p,~MU_p,~JJ+Met_eff,grue.data)
MU_p_MHo_p_MU_p<-colext(~MU_p,~MHo_p,~MU_p,~JJ+Met_eff,grue.data)
MAt_sMAt_2_MHo_p_MU_p<-colext(~MAt_s+MAt_2,~MHo_p,~MU_p,~JJ+Met_eff,grue.data)
MAi_sMAi_2_MHo_p_MU_p<-colext(~MAi_s+MAi_2,~MHo_p,~MU_p,~JJ+Met_eff,grue.data)
MHo_p_MU_p_MU_p<-colext(~MHo_p,~MU_p,~MU_p,~JJ+Met_eff,grue.data)
MU_p_MU_p_MU_p<-colext(~MU_p,~MU_p,~MU_p,~JJ+Met_eff,grue.data)
MAt_sMAt_2_MU_p_MU_p<-colext(~MAt_s+MAt_2,~MU_p,~MU_p,~JJ+Met_eff,grue.data)
MAi_sMAi_2_MU_p_MU_p<-colext(~MAi_s+MAi_2,~MU_p,~MU_p,~JJ+Met_eff,grue.data)
MHo_p_MAt.sMAt2_MU_p<-colext(~MHo_p,~MAt.s+MAt2,~MU_p,~JJ+Met_eff,grue.data)
MU_p_MAt.sMAt2_MU_p<-colext(~MU_p,~MAt.s+MAt2,~MU_p,~JJ+Met_eff,grue.data)
MAt_sMAt_2_MAt.sMAt2_MU_p<-colext(~MAt_s+MAt_2,~MAt.s+MAt2,~MU_p,~JJ+Met_eff,grue.data)
MAi_sMAi_2_MAt.sMAt2_MU_p<-colext(~MAi_s+MAi_2,~MAt.s+MAt2,~MU_p,~JJ+Met_eff,grue.data)
MHo_p_MAi.sMAi2_MU_p<-colext(~MHo_p,~MAi.s+MAi2,~MU_p,~JJ+Met_eff,grue.data)
MU_p_MAi.sMAi2_MU_p<-colext(~MU_p,~MAi.s+MAi2,~MU_p,~JJ+Met_eff,grue.data)
MAt_sMAt_2_MAi.sMAi2_MU_p<-colext(~MAt_s+MAt_2,~MAi.s+MAi2,~MU_p,~JJ+Met_eff,grue.data)
MAi_sMAi_2_MAi.sMAi2_MU_p<-colext(~MAi_s+MAi_2,~MAi.s+MAi2,~MU_p,~JJ+Met_eff,grue.data)
MHo_p_MHo_p_MAt.sMAt2<-colext(~MHo_p,~MHo_p,~MAt.s+MAt2,~JJ+Met_eff,grue.data)
MU_p_MHo_p_MAt.sMAt2<-colext(~MU_p,~MHo_p,~MAt.s+MAt2,~JJ+Met_eff,grue.data)
MAt_sMAt_2_MHo_p_MAt.sMAt2<-colext(~MAt_s+MAt_2,~MHo_p,~MAt.s+MAt2,~JJ+Met_eff,grue.data)
MAi_sMAi_2_MHo_p_MAt.sMAt2<-colext(~MAi_s+MAi_2,~MHo_p,~MAt.s+MAt2,~JJ+Met_eff,grue.data)
MHo_p_MU_p_MAt.sMAt2<-colext(~MHo_p,~MU_p,~MAt.s+MAt2,~JJ+Met_eff,grue.data)
MU_p_MU_p_MAt.sMAt2<-colext(~MU_p,~MU_p,~MAt.s+MAt2,~JJ+Met_eff,grue.data)
MAt_sMAt_2_MU_p_MAt.sMAt2<-colext(~MAt_s+MAt_2,~MU_p,~MAt.s+MAt2,~JJ+Met_eff,grue.data)
MAi_sMAi_2_MU_p_MAt.sMAt2<-colext(~MAi_s+MAi_2,~MU_p,~MAt.s+MAt2,~JJ+Met_eff,grue.data)
MHo_p_MAt.sMAt2_MAt.sMAt2<-colext(~MHo_p,~MAt.s+MAt2,~MAt.s+MAt2,~JJ+Met_eff,grue.data)
MU_p_MAt.sMAt2_MAt.sMAt2<-colext(~MU_p,~MAt.s+MAt2,~MAt.s+MAt2,~JJ+Met_eff,grue.data)
MAt_sMAt_2_MAt.sMAt2_MAt.sMAt2<-colext(~MAt_s+MAt_2,~MAt.s+MAt2,~MAt.s+MAt2,~JJ+Met_eff,grue.data)
MAi_sMAi_2_MAt.sMAt2_MAt.sMAt2<-colext(~MAi_s+MAi_2,~MAt.s+MAt2,~MAt.s+MAt2,~JJ+Met_eff,grue.data)
MHo_p_MAi.sMAi2_MAt.sMAt2<-colext(~MHo_p,~MAi.s+MAi2,~MAt.s+MAt2,~JJ+Met_eff,grue.data)
MU_p_MAi.sMAi2_MAt.sMAt2<-colext(~MU_p,~MAi.s+MAi2,~MAt.s+MAt2,~JJ+Met_eff,grue.data)
MAt_sMAt_2_MAi.sMAi2_MAt.sMAt2<-colext(~MAt_s+MAt_2,~MAi.s+MAi2,~MAt.s+MAt2,~JJ+Met_eff,grue.data)
MAi_sMAi_2_MAi.sMAi2_MAt.sMAt2<-colext(~MAi_s+MAi_2,~MAi.s+MAi2,~MAt.s+MAt2,~JJ+Met_eff,grue.data)
MHo_p_MHo_p_MAi.sMAi2<-colext(~MHo_p,~MHo_p,~MAi.s+MAi2,~JJ+Met_eff,grue.data)
MU_p_MHo_p_MAi.sMAi2<-colext(~MU_p,~MHo_p,~MAi.s+MAi2,~JJ+Met_eff,grue.data)
MAt_sMAt_2_MHo_p_MAi.sMAi2<-colext(~MAt_s+MAt_2,~MHo_p,~MAi.s+MAi2,~JJ+Met_eff,grue.data)
MAi_sMAi_2_MHo_p_MAi.sMAi2<-colext(~MAi_s+MAi_2,~MHo_p,~MAi.s+MAi2,~JJ+Met_eff,grue.data)
MHo_p_MU_p_MAi.sMAi2<-colext(~MHo_p,~MU_p,~MAi.s+MAi2,~JJ+Met_eff,grue.data)
MU_p_MU_p_MAi.sMAi2<-colext(~MU_p,~MU_p,~MAi.s+MAi2,~JJ+Met_eff,grue.data)
MAt_sMAt_2_MU_p_MAi.sMAi2<
  -colext(~MAt_s+MAt_2,~MU_p,~MAi.s+MAi2,~JJ+Met_eff,grue.data)
MAi_sMAi_2_MU_p_MAi.sMAi2<-colext(~MAi_s+MAi_2,~MU_p,~MAi.s+MAi2,~JJ+Met_eff,grue.data)
MHo_p_MAt.sMAt2_MAi.sMAi2<-colext(~MHo_p,~MAt.s+MAt2,~MAi.s+MAi2,~JJ+Met_eff,grue.data)
MU_p_MAt.sMAt2_MAi.sMAi2<-colext(~MU_p,~MAt.s+MAt2,~MAi.s+MAi2,~JJ+Met_eff,grue.data)
MAt_sMAt_2_MAt.sMAt2_MAi.sMAi2<-colext(~MAt_s+MAt_2,~MAt.s+MAt2,~MAi.s+MAi2,~JJ+Met_eff,grue.data)
MAi_sMAi_2_MAt.sMAt2_MAi.sMAi2<-colext(~MAi_s+MAi_2,~MAt.s+MAt2,~MAi.s+MAi2,~JJ+Met_eff,grue.data)
MHo_p_MAi.sMAi2_MAi.sMAi2<-colext(~MHo_p,~MAi.s+MAi2,~MAi.s+MAi2,~JJ+Met_eff,grue.data)
MU_p_MAi.sMAi2_MAi.sMAi2<-colext(~MU_p,~MAi.s+MAi2,~MAi.s+MAi2,~JJ+Met_eff,grue.data)
MAt_sMAt_2_MAi.sMAi2_MAi.sMAi2<-colext(~MAt_s+MAt_2,~MAi.s+MAi2,~MAi.s+MAi2,~JJ+Met_eff,grue.data)
MAi_sMAi_2_MAi.sMAi2_MAi.sMAi2<-colext(~MAi_s+MAi_2,~MAi.s+MAi2,~MAi.s+MAi2,~JJ+Met_eff,grue.data)

hab_dyn <- c(MHo_p_MHo_p_MHo_p,MU_p_MHo_p_MHo_p,MAt_sMAt_2_MHo_p_MHo_p,MAi_sMAi_2_MHo_p_MHo_p,MHo_p_MU_p_MHo_p,MU_p_MU_p_MHo_p,MAt_sMAt_2_MU_p_MHo_p,MAi_sMAi_2_MU_p_MHo_p,MHo_p_MAt.sMAt2_MHo_p,MU_p_MAt.sMAt2_MHo_p,MAt_sMAt_2_MAt.sMAt2_MHo_p,MAi_sMAi_2_MAt.sMAt2_MHo_p,MHo_p_MAi.sMAi2_MHo_p,MU_p_MAi.sMAi2_MHo_p,MAt_sMAt_2_MAi.sMAi2_MHo_p,MAi_sMAi_2_MAi.sMAi2_MHo_p,MHo_p_MHo_p_MU_p,MU_p_MHo_p_MU_p,MAt_sMAt_2_MHo_p_MU_p,MAi_sMAi_2_MHo_p_MU_p,MHo_p_MU_p_MU_p,MU_p_MU_p_MU_p,MAt_sMAt_2_MU_p_MU_p,MAi_sMAi_2_MU_p_MU_p,MHo_p_MAt.sMAt2_MU_p,MU_p_MAt.sMAt2_MU_p,MAt_sMAt_2_MAt.sMAt2_MU_p, MAi_sMAi_2_MAt.sMAt2_MU_p,MHo_p_MAi.sMAi2_MU_p,MU_p_MAi.sMAi2_MU_p,MAt_sMAt_2_MAi.sMAi2_MU_p,MAi_sMAi_2_MAi.sMAi2_MU_p,MHo_p_MHo_p_MAt.sMAt2,MU_p_MHo_p_MAt.sMAt2,MAt_sMAt_2_MHo_p_MAt.sMAt2,MAi_sMAi_2_MHo_p_MAt.sMAt2,MHo_p_MU_p_MAt.sMAt2,MU_p_MU_p_MAt.sMAt2,MAt_sMAt_2_MU_p_MAt.sMAt2,MAi_sMAi_2_MU_p_MAt.sMAt2,MHo_p_MAt.sMAt2_MAt.sMAt2,MU_p_MAt.sMAt2_MAt.sMAt2,MAt_sMAt_2_MAt.sMAt2_MAt.sMAt2,MAi_sMAi_2_MAt.sMAt2_MAt.sMAt2,MHo_p_MAi.sMAi2_MAt.sMAt2,MU_p_MAi.sMAi2_MAt.sMAt2,MAt_sMAt_2_MAi.sMAi2_MAt.sMAt2,MAi_sMAi_2_MAi.sMAi2_MAt.sMAt2,MHo_p_MHo_p_MAi.sMAi2,MU_p_MHo_p_MAi.sMAi2,MAt_sMAt_2_MHo_p_MAi.sMAi2,MAi_sMAi_2_MHo_p_MAi.sMAi2,MHo_p_MU_p_MAi.sMAi2,MU_p_MU_p_MAi.sMAi2,MAt_sMAt_2_MU_p_MAi.sMAi2,MAi_sMAi_2_MU_p_MAi.sMAi2,MHo_p_MAt.sMAt2_MAi.sMAi2,MU_p_MAt.sMAt2_MAi.sMAi2,MAt_sMAt_2_MAt.sMAt2_MAi.sMAi2,MAi_sMAi_2_MAt.sMAt2_MAi.sMAi2,MHo_p_MAi.sMAi2_MAi.sMAi2,MU_p_MAi.sMAi2_MAi.sMAi2,MAt_sMAt_2_MAi.sMAi2_MAi.sMAi2 ,MAi_sMAi_2_MAi.sMAi2_MAi.sMAi2)

names(hab_dyn)<- c("MHo_p_MHo_p_MHo_p","MU_p_MHo_p_MHo_p","MAt_sMAt_2_MHo_p_MHo_p","MAi_sMAi_2_MHo_p_MHo_p","MHo_p_MU_p_MHo_p"," MU_p_MU_p_MHo_p","MAt_sMAt_2_MU_p_MHo_p","MAi_sMAi_2_MU_p_MHo_p","MHo_p_MAt.sMAt2_MHo_p","MU_p_MAt.sMAt2_MHo_p","MAt_sMAt_2_MAt.sMAt2_MHo_p","MAi_sMAi_2_MAt.sMAt2_MHo_p","MHo_p_MAi.sMAi2_MHo_p","MU_p_MAi.sMAi2_MHo_p","MAt_sMAt_2_MAi.sMAi2_MHo_p","MAi_sMAi_2_MAi.sMAi2_MHo_p","MHo_p_MHo_p_MU_p","MU_p_MHo_p_MU_p","MAt_sMAt_2_MHo_p_MU_p","MAi_sMAi_2_MHo_p_MU_p","MHo_p_MU_p_MU_p","MU_p_MU_p_MU_p","MAt_sMAt_2_MU_p_MU_p","MAi_sMAi_2_MU_p_MU_p"," MHo_p_MAt.sMAt2_MU_p","MU_p_MAt.sMAt2_MU_p","MAt_sMAt_2_MAt.sMAt2_MU_p","MAi_sMAi_2_MAt.sMAt2_MU_p","MHo_p_MAi.sMAi2_MU_p","MU_p_MAi.sMAi2_MU_p","MAt_sMAt_2_MAi.sMAi2_MU_p","MAi_sMAi_2_MAi.sMAi2_MU_p","MHo_p_MHo_p_MAt.sMAt2","MU_p_MHo_p_MAt.sMAt2","MAt_sMAt_2_MHo_p_MAt.sMAt2","MAi_sMAi_2_MHo_p_MAt.sMAt2","MHo_p_MU_p_MAt.sMAt2","MU_p_MU_p_MAt.sMAt2","MAt_sMAt_2_MU_p_MAt.sMAt2","MAi_sMAi_2_MU_p_MAt.sMAt2","MHo_p_MAt.sMAt2_MAt.sMAt2","MU_p_MAt.sMAt2_MAt.sMAt2","MAt_sMAt_2_MAt.sMAt2_MAt.sMAt2","MAi_sMAi_2_MAt.sMAt2_MAt.sMAt2","MHo_p_MAi.sMAi2_MAt.sMAt2","MU_p_MAi.sMAi2_MAt.sMAt2","MAt_sMAt_2_MAi.sMAi2_MAt.sMAt2","MAi_sMAi_2_MAi.sMAi2_MAt.sMAt2","MHo_p_MHo_p_MAi.sMAi2","MU_p_MHo_p_MAi.sMAi2","MAt_sMAt_2_MHo_p_MAi.sMAi2","MAi_sMAi_2_MHo_p_MAi.sMAi2","MHo_p_MU_p_MAi.sMAi2","MU_p_MU_p_MAi.sMAi2","MAt_sMAt_2_MU_p_MAi.sMAi2","MAi_sMAi_2_MU_p_MAi.sMAi2","MHo_p_MAt.sMAt2_MAi.sMAi2","MU_p_MAt.sMAt2_MAi.sMAi2","MAt_sMAt_2_MAt.sMAt2_MAi.sMAi2","MAi_sMAi_2_MAt.sMAt2_MAi.sMAi2","MHo_p_MAi.sMAi2_MAi.sMAi2","MU_p_MAi.sMAi2_MAi.sMAi2","MAt_sMAt_2_MAi.sMAi2_MAi.sMAi2","MAi_sMAi_2_MAi.sMAi2_MAi.sMAi2")



#load("hab_dyn2.Rdata")

```

Test GOF (Running this take several hours) - You have to activated to run using eval = TRUE, include = TRUE
```{r eval=FALSE, , eval=FALSE, include=FALSE}
#Test de MacKenzie et Bailey for checking model adjustment
#gof <- mb.gof.test(MHo_p_MAt.sMAt2_MHo_p, nsim = 1000, plot.hist = TRUE)
#save(gof, file = "MHo_p_MAt.sMAt2_MHo_p_dyn_gof.Rdata")

gof <- gof6 # If you load our true gof, you will see the plot.


## C.hat
gof$chisq.table
gof$c.hat.est


# Plots
hist(gof$t.star, 
     main = as.expression(substitute("Bootstrapped "*chi^2*" fit statistic ("*nsim*" samples)",
         list(nsim = gof$nsim))), 
     xlim = range(c(gof$t.star, gof$chi.square)), 
     xlab = paste("Simulated statistic ", 
       "(observed = ", round(gof$chi.square, digits = 2), ")", 
       sep = ""))
title(main = bquote(paste(italic(P), " = ", .(gof$p.global))), 
      line = -0.5)
abline(v = gof$chi.square, lty = "dashed", col = "red")


```

AIC selection for habitat models (you need to run the models before)

```{r selection modeles, eval=FALSE, include=FALSE}

aictab(cand.set = hab_dyn,c.hat = 1)

# check parametres

evidence(aictab(cand.set = hab_dyn, c.hat = 1))
lapply(hab_dyn, extractCN)
lapply(hab_dyn, checkParms, simplify = FALSE)

## checker convergence des modeles
lapply(hab_dyn, checkConv) 

summary(MHo_p_MAt.sMAt2_MHo_p)

```

Habitat figures 

```{r figures habitat, echo=FALSE}

# New DB for wetlands 

MHo_p.mean <- mean(SiteCovs$MHo_p)
MHo_p.sd <- sd(SiteCovs$MHo_p)

mh <- data.frame(MHo = seq(from=min(SiteCovs$MHo_p, na.rm = TRUE),
                           to = max(SiteCovs$MHo_p,na.rm = TRUE),
                           length=50),
                       MU_p= 0,
                       MAi = 0,
                       MAt = 0,
                       JJ = 0,
                       km = 0,
                 mm = 0,
                 temp = 0,
                 An = 0,
                       Met_eff = factor("HHT",levels =levels(grue.data@obsCovs$Met_eff)))

mh$MHo_p <- (mh$MHo-MHo_p.mean)/MHo_p.sd 

                        
E.psi <- predict(MHo_p_MAt.sMAt2_MHo_p, type="psi", newdata=mh, appendData=TRUE)
E.psi$fit <- E.psi$Predicted
E.psi$se.fit <- E.psi$SE
E.psi$lwl <- E.psi$fit-1.96*E.psi$se.fit
E.psi$upl <- E.psi$fit+1.96*E.psi$se.fit
 
 Plot_psi_openwetlands <- E.psi %>%
  ggplot(aes(MHo, fit)) +
  geom_smooth_ci(color = "black") +
  theme_bw() +
  ylim(-0.02,1.05) +
  xlim(0,40) +
  geom_line(linetype = "dotted", aes(y = lwl), color = "grey43") +
  geom_line(linetype = "dotted", aes(y = upl), color = "grey43") +
  xlab("% Open wetlands") + 
  ylab(expression(paste('Initial occupancy ( ', psi, ' )'))) +
  theme(
      axis.title.x = element_text(family="Times New Roman", color="black", size = 16),
      axis.title.y = element_text(family="Times New Roman",color="black", size=16),
      axis.text=element_text(family="Times New Roman",size=12)
    )
 

E.ext <- predict(MHo_p_MAt.sMAt2_MHo_p, type="ext", newdata=mh, appendData=TRUE)
E.ext$fit <- E.ext$Predicted
E.ext$se.fit <- E.ext$SE
E.ext$lwl <- E.ext$fit-1.96*E.ext$se.fit
E.ext$upl <- E.ext$fit+1.96*E.ext$se.fit
 
 Plot_extinction_openwetlands <- E.ext %>%
  ggplot(aes(MHo, fit)) +
  geom_smooth_ci(color = "black") +
  theme_bw() +
  ylim(-0.02,1) +
   xlim(0,40) +
  geom_line(linetype = "dotted", aes(y = lwl), color = "grey43") +
  geom_line(linetype = "dotted", aes(y = upl), color = "grey43") +
  xlab("% Open wetlands") + 
  ylab(expression(paste('Extinction probability ( ', epsilon, ' )'))) +
  theme(
      axis.title.x = element_text(family="Times New Roman", color="black", size = 16),
      axis.title.y = element_text(family="Times New Roman",color="black", size=16),
      axis.text=element_text(family="Times New Roman",size=12)
    )



# Variables qu'influencent l'extinction

ma <- data.frame(MAt = seq(from=min(MAt),
                           to = max(MAt),
                           length=50),
                       MHo_p= 0,
                       MAi = 0,
                       MU_p = 0,
                       JJ = 0,
                       km = 0,
                 mm = 0,
                 temp = 0,
                 An = 0,
                       Met_eff = factor("HHT",levels =levels(grue.data@obsCovs$Met_eff)))
ma$MAt.s <-(ma$MAt-MAt.mean)/MAt.sd 
ma$MAt2 <- ma$MAt.s^2
E.col<- predict(MHo_p_MAt.sMAt2_MHo_p, type="col", newdata=ma, appendData=TRUE)
E.col$fit <- E.col$Predicted
E.col$se.fit <- E.col$SE
E.col$lwl <- E.col$fit-1.96*E.col$se.fit
E.col$upl <- E.col$fit+1.96*E.col$se.fit

Plot_col_agricultural <- E.col %>%
  ggplot(aes(MAt, fit)) +
  geom_smooth_ci(color = "black") +
  theme_bw() +
  ylim(0,1) +
  xlim(0,90)+
  geom_line(linetype = "dotted", aes(y = lwl), color = "grey43") +
  geom_line(linetype = "dotted", aes(y = upl), color = "grey43") +
  xlab("% Agricultural landscape") + 
  ylab(expression(paste('Colonization probability ( ', gamma, ' )'))) +
  theme(
      axis.title.x = element_text(family="Times New Roman", color="black", size = 16),
      axis.title.y = element_text(family="Times New Roman",color="black", size=16),
      axis.text=element_text(family="Times New Roman",size=12)
    )

Plot_psi_openwetlands
Plot_extinction_openwetlands
Plot_col_agricultural


```

